---
title       : Exploring Normal Mixtures
subtitle    : A Prototype
author      : Xan Gregg
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Up Front

* The Old Faithful geyser eruption is famously bimodal, but is each mode normal?
* For other data sets, it can be useful to try a normal mixture distribution.
* This app is limited to the geyser data and serves as a prototype for more elaborate exploration.

---

## Computation

The mixtools package is used for the normal mixture solution. This example shows the results of fitting a mixture of two normals.

```{r}
 library(mixtools)
 fit <- normalmixEM(faithful$eruptions, k=2);
 show(fit$mu);show(fit$sigma)
```

---


## Graph Output

Individual normal curves are plotted over a histogram of the data.
```{r echo=FALSE}
 library(mixtools)
 fit2 <- normalmixEM(faithful$eruptions, k=2);
  hist(faithful$eruptions,  probability = TRUE, ylim=c(0,1.2),
    breaks = as.numeric(30), xlab = "Duration (minutes)", main = "")

    for(i in 1:2) { curve(dnorm(x,mean=fit2$mu[i], sd=fit2$sigma[i])/2, col=i+1, lwd=2, add=TRUE) }
```

---


## Next Steps

* Support arbitrary data sets
* Plot the combined mixture distribution
* Show the paramaters for each fit (mean and standard deviation)
* Show statistics that quantify the quality of the fit





